---
title: "Data exploration2"
author: "Mengfan Luo"
date: "11/11/2021"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(httr)
library(haven)
```


## Data loading and crude cleaning

We select year 2014-2016 and variables relating to smoking and insurance. 

```{r}

chs16 = read_sas("data/chs2016_public.sas7bdat")

chs16_filter = chs16 %>% 
  select(agegroup,generalhealth,insuredgateway16,insure16,insured,insure5,sickadvice16,sickplace,didntgetcare16,smoker,everyday,numberperdaya,cost20cigarettes,imputed_povertygroup ,bmi,child,sex) %>% 
  mutate(year = 2016) %>% 
  rename(insuredgateway = insuredgateway16, insure = insure16,sickadvice = sickadvice16,didntgetcare = didntgetcare16)


chs15 = read_sas("data/chs2015_public.sas7bdat")

chs15_filter = chs15 %>% 
  select(agegroup,generalhealth,insuredgateway15,insure15,insured,insure5,sickadvice15,sickplace,didntgetcare15,smoker,everyday,numberperdaya,cost20cigarettes,imputed_povertygroup ,bmi,child,sex) %>% 
  mutate(year = 2015) %>% 
  rename(insuredgateway = insuredgateway15, insure = insure15,sickadvice = sickadvice15,didntgetcare = didntgetcare15)


chs14 = read_sas("data/chs2014_public.sas7bdat")

chs14_filter = chs14 %>% 
  select(agegroup,generalhealth,insuredgateway14,insure14,insured,insure5,sickadvice14,sickplace,didntgetcare14,smoker,everyday,numberperdaya,cost20cigarettes,imputed_povertygroup ,bmi,child,sex) %>% 
  mutate(year = 2014) %>% 
  rename(insuredgateway = insuredgateway14, insure = insure14,sickadvice = sickadvice14,didntgetcare = didntgetcare14)


chs_14_16 = bind_rows(chs14_filter,chs15_filter,chs16_filter)
  
```


```{r}
dataset_basic = chs_14_16 %>% 
  select(agegroup,insuredgateway,insure,sickadvice,sickadvice,didntgetcare,smoker,everyday,numberperdaya,cost20cigarettes,generalhealth,imputed_povertygroup,bmi,child,sex) %>% 
  mutate(numberperdaya = round(numberperdaya,1),
         imputed_povertygroup = factor(imputed_povertygroup),
         child = factor(child,levels = c(1,2),labels = c(1,0)),
         sex = factor(sex))
write_csv(dataset_basic,"data/dataset_update.csv")
```


## About smoking

```{r}

smoke_df = dataset_basic %>% 
  select(agegroup,smoker,everyday,numberperdaya,cost20cigarettes,generalhealth,imputed_povertygroup,bmi,child,sex) %>% 
  drop_na(smoker) %>% 
  mutate(smoker = factor(smoker,levels = c(1,2,3),labels = c("never","current","former")),
         everyday = factor(everyday),
         agegroup = factor(agegroup,ordered = TRUE,labels = c("18-24yrs","25-44yrs", "45-64yrs", "65+yrs")),
         smoker_cat = case_when(
           numberperdaya <= 10 ~ 1,
           (numberperdaya > 10) & (numberperdaya <= 20) ~2,
           numberperdaya > 20 ~ 3),
         smoker_cat = factor(smoker_cat,levels = c(1,2,3),labels = c("light","normal","heavy")),
         generalhealth = factor(generalhealth,levels = c(5,4,3,2,1), labels = c("Poor","Fair","Good","Very good","Excellent"))
         
         )
```



* Age distribution among people in different smoking status

```{r}

smoke_df %>% 
  drop_na(agegroup) %>% 
  group_by(agegroup,smoker) %>% 
  summarize(
    count = n(),
  ) %>% 
  mutate(smoker = fct_reorder(smoker,count)) %>% 
  ggplot(aes(x = smoker, y = count, fill = agegroup)) + 
  geom_bar(stat = "identity",position = "dodge") +
    labs(
      x = 'Smoking status',
      y = 'Number of smokers',
      title = 'Age distribution among people in different smoking status')


```


* Self-evaluation of health conditions among people in different smoking status

```{r}

smokegroup_count = smoke_df %>% 
  drop_na(smoker) %>% 
  group_by(smoker) %>% 
  summarize(
    smokegroup_count = n()) %>% 
  pull()

smoke_df %>% 
  drop_na(generalhealth) %>% 
  group_by(generalhealth,smoker) %>% 
  summarize(
    count = n())%>% 
  mutate(smokegroup_count = smokegroup_count,
         health_percent = count/smokegroup_count)%>% 
  ggplot(aes(x = smoker, y = health_percent, fill = generalhealth)) + 
  geom_bar(stat = "identity",position = "dodge") +
    labs(
      x = 'Type of smokers',
      y = 'Health self-evaluation \n(% in each types of smoker)',
      title = 'Self-evaluation of health conditions among people in different smoking status\n(% in each types of smoker)')
```


```{r}
smoke_df %>% 
  drop_na(imputed_povertygroup) %>% 
  group_by(imputed_povertygroup,smoker) %>% 
  summarize(
    count = n(),
  ) %>% 
  mutate(
    smokegroup_count = smokegroup_count,
    poverty_percent = count/smokegroup_count,
    smoker = fct_reorder(smoker,count)) %>% 
  ggplot(aes(x = smoker, y = poverty_percent, fill = imputed_povertygroup)) + 
  geom_bar(stat = "identity",position = "dodge") +
    labs(
      x = 'Smoking status',
      y = 'imputed_povertygroup',
      title = 'povertygroup among different smoking status')
```



For `never` and `former` smokers, no data for variables `everyday`, `numberperdaya`, `cost20cigarettes`, so next we focus on the current smokers.


* Cost and number of daily consumption

```{r}
current_smoker = smoke_df %>% filter(smoker == "current")

current_smoker %>%
  ggplot(aes(x = smoker_cat, y = cost20cigarettes, fill = smoker_cat))+
  geom_violin() +
    labs(
      x = 'Type of smokers',
      y = 'Cost of every 20 (one pack) cigrattes',
      title = 'Cost of every 20 (one pack) cigrattes\namong different type of smokers')
  
  
current_smoker %>% 
  ggplot(aes(x = numberperdaya, y = cost20cigarettes,color = smoker_cat))+
  geom_point(alpha = .5,size = 3) +
  facet_grid(.~smoker_cat,scales = "free") +
    labs(
      x = 'Number of cigrattes consumed everyday',
      y = 'Cost of every 20 (one pack) cigrattes',
      title = 'Cost of every 20 (one pack) cigrattes\nfor number of cigrattes consumed everyday')


```


Chi-squared test for `insure` and 

2.

```{r}
analyse_poverty_insurancetype = 
  dataset_basic %>%
  drop_na(imputed_povertygroup,insure) %>% 
  mutate(imputed_povertygroup = factor(imputed_povertygroup,levels = c(1,2,3,4,5))) %>% 
  mutate(insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid/Family Health+", "Milit/CHAMPUS/Tricare", "COBRA/Other", "Uninsured"))) %>% 
  group_by(imputed_povertygroup,insure) %>% 
  summarize(n = n())
```


```{r}
test_data_poverty_insurance = analyse_poverty_insurancetype %>% 
  pivot_wider(
  names_from = "insure", 
  values_from = "n")
```

```{r}
chisq.test(test_data_poverty_insurance[-1])

```







