---
title: "Data Exploaration"
output: 
  html_document:
    code_folding: hide
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(httr)
library(haven)
library(patchwork)
library(readr)
```


## Data loading and crude cleaning

We select year 2014-2016 and variables relating to smoking and insurance. 

```{r}

chs16 = read_sas("data/chs2016_public.sas7bdat")

chs16_filter = chs16 %>% 
  select(agegroup,generalhealth,insuredgateway16,insure16,insured,insure5,sickadvice16,sickplace,didntgetcare16,smoker,everyday,numberperdaya,cost20cigarettes,imputed_povertygroup,fluvaccineshot) %>% 
  mutate(year = 2016) %>% 
  rename(insuredgateway = insuredgateway16, insure = insure16,sickadvice = sickadvice16,didntgetcare = didntgetcare16)


chs15 = read_sas("data/chs2015_public.sas7bdat")

chs15_filter = chs15 %>% 
  select(agegroup,generalhealth,insuredgateway15,insure15,insured,insure5,sickadvice15,sickplace,didntgetcare15,smoker,everyday,numberperdaya,cost20cigarettes,imputed_povertygroup,fluvaccineshot) %>% 
  mutate(year = 2015) %>% 
  rename(insuredgateway = insuredgateway15, insure = insure15,sickadvice = sickadvice15,didntgetcare = didntgetcare15)


chs14 = read_sas("data/chs2014_public.sas7bdat")

chs14_filter = chs14 %>% 
  select(agegroup,generalhealth,insuredgateway14,insure14,insured,insure5,sickadvice14,sickplace,didntgetcare14,smoker,everyday,numberperdaya,cost20cigarettes,imputed_povertygroup,fluvaccineshot) %>% 
  mutate(year = 2014) %>% 
  rename(insuredgateway = insuredgateway14, insure = insure14,sickadvice = sickadvice14,didntgetcare = didntgetcare14)


chs_14_16 = bind_rows(chs14_filter,chs15_filter,chs16_filter)
  





```


```{r}
dataset_basic = chs_14_16 %>% 
  select(agegroup,insuredgateway,insure,sickadvice,sickadvice,didntgetcare,smoker,everyday,numberperdaya,cost20cigarettes,generalhealth,fluvaccineshot,imputed_povertygroup) %>% 
  mutate(numberperdaya = round(numberperdaya,1))

write_csv(dataset_basic,"data/dataset_basic.csv")

```


```{r}
insurance_data = dataset_basic %>% 
  select(insuredgateway,insure,sickadvice,didntgetcare)


plot1 = insurance_data %>% 
  ggplot(aes(x = insuredgateway)) +
  geom_histogram() 
 
plot2 =insurance_data %>% 
  ggplot(aes(x = insure)) +
  geom_histogram()  

plot3 =insurance_data %>% 
  ggplot(aes(x = sickadvice)) +
  geom_histogram() 

plot4 =insurance_data %>% 
  ggplot(aes(x = didntgetcare)) +
  geom_histogram() 

plot1 + plot2 + plot3 + plot4
```




### Interested variables and smoking {.tabset}

#### insure

```{r}
analyse_smoke_insurancetype = 
  dataset_basic %>%
  drop_na(smoker,insure) %>% 
  mutate(smoker = factor(smoker,levels = c(1,2,3),labels = c("never","current","former")),
         everyday = factor(everyday)) %>% 
  mutate(insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid/Family Health+", "Milit/CHAMPUS/Tricare", "COBRA/Other", "Uninsured"))) %>% 
  group_by(smoker,insure) %>% 
  summarize(n = n())
  
analyse_smoke_insurancetype %>% 
  ggplot(aes(x = insure, y = n, fill = smoker)) + 
  geom_bar(stat = "identity",position = "dodge") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) +
    labs(
      x = 'insure type',
      y = 'counts',
      title = 'The distribution of insure in different smoking group ')
```

```{r}
test_data_smoke_insurancetype = analyse_smoke_insurancetype %>% 
  pivot_wider(
  names_from = "insure", 
  values_from = "n")
```

```{r}
chisq.test(test_data_smoke_insurancetype[-1])

```

#### healthstatus

```{r}
analyse_generalhealth_insurancetype = 
  dataset_basic %>%
  drop_na(generalhealth,insure) %>% 
  mutate(insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid/Family Health+", "Milit/CHAMPUS/Tricare", "COBRA/Other", "Uninsured")))%>% 
  mutate(generalhealth = factor(generalhealth,levels = c(1,2,3,4,5),labels = c("Excellent","Very good","Good", "Fair", "Poor"))) %>% 
  group_by(insure,generalhealth) %>% 
  summarize(n = n())
  
analyse_generalhealth_insurancetype %>% 
  ggplot(aes(x = insure, y = n, fill = generalhealth )) + 
  geom_bar(stat = "identity",position = "dodge") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) +
    labs(
      x = 'generalhealth',
      y = 'counts',
      title = 'The distribution of generalhealth in different smoking group ')
```

```{r}
test_data_generalhealth_insurancetype = analyse_generalhealth_insurancetype %>% 
  pivot_wider(
  names_from = "insure", 
  values_from = "n")
```

```{r}
test_generalhealth = chisq.test(test_data_generalhealth_insurancetype[-1])

```

#### when sick, where usually go?

```{r}
analyse_sickadvice_insurancetype = 
  dataset_basic %>%
  drop_na(sickadvice,insure) %>% 
  mutate(insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid_Family Health", "milit_champus_tricare", "COBRA_Other", "uninsured")))%>% 
  mutate(sickadvice = factor(sickadvice,levels = c(1,2,3,4,5,6,7,8,9),labels = c("private_doctor","non-retail_clinic","urgent_care_center", "hospital_ed", "retail_clinic", "alternative_health_care_provider","family_friend_self_resources","other","no_usual_place"))) %>% 
  group_by(insure,sickadvice) %>% 
  summarize(n = n())
  
analyse_sickadvice_insurancetype %>% 
  ggplot(aes(x = insure, y = n, fill = sickadvice )) + 
  geom_bar(stat = "identity",position = "dodge") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) +
    labs(
      x = 'sickadvice',
      y = 'counts',
      title = 'The distribution of sickadvice in different smoking group ')
```

```{r}
test_data_sickadvice_insurancetype = analyse_sickadvice_insurancetype %>% 
  pivot_wider(
  names_from = "insure", 
  values_from = "n") %>% 
  mutate(milit_champus_tricare = replace_na(milit_champus_tricare, 0))
```

```{r}
test_generalhealth = chisq.test(test_data_sickadvice_insurancetype[-1])

```
According to above chi-square test result We reject the null hypothesis and conclude that there’s at least one helath stutus group’s proportions of insurance type is different from others at 0.05 significant level.


#### Was there a time in the past 12 months when you needed medical care but did NOT get it?

```{r}
analyse_didntgetcare_insurancetype = 
  dataset_basic %>%
  drop_na(didntgetcare,insure) %>% 
  mutate(insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid_Family Health", "milit_champus_tricare", "COBRA_Other", "uninsured")))%>% 
  mutate(didntgetcare = factor(didntgetcare,levels = c(1,2),labels = c("Yes","No"))) %>% 
  group_by(insure,didntgetcare) %>% 
  summarize(n = n())
  
analyse_didntgetcare_insurancetype %>% 
  ggplot(aes(x = insure, y = n, fill = didntgetcare )) + 
  geom_bar(stat = "identity",position = "dodge") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) +
    labs(
      x = 'didntgetcare',
      y = 'counts',
      title = 'The distribution of didntgetcare in different smoking group ')
```

```{r}
test_data_didntgetcare_insurancetype = analyse_didntgetcare_insurancetype %>% 
  pivot_wider(
  names_from = "insure", 
  values_from = "n") 
  #mutate(milit_champus_tricare = replace_na(milit_champus_tricare, 0))
```

```{r}
test_generalhealth = chisq.test(test_data_didntgetcare_insurancetype[-1])

```