---
title: "data exploration"
output: 
  html_document:
    code_folding: hide
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(httr)
library(haven)
library(patchwork)
library(summarytools)
library(grid)
library(vcd)

```

```{r message=FALSE}
dataset_basic = read_csv("data/dataset_basic.csv")
```


### Variables and insurance {.tabset}

#### Smoking

##### Plot

```{r message=FALSE}
analyse_smoke_insurancetype = 
  dataset_basic %>%
  drop_na(smoker,insure) %>% 
  mutate(smoker = factor(smoker,levels = c(1,2,3),labels = c("never","current","former")),
         everyday = factor(everyday)) %>% 
  mutate(insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid/Family Health+", "Milit/CHAMPUS/Tricare", "COBRA/Other", "Uninsured"))) %>% 
  group_by(smoker,insure) %>% 
  summarize(n = n())
  
analyse_smoke_insurancetype %>% 
  ggplot(aes(x =smoker , y = n, fill = insure)) + 
  geom_bar(stat = "identity",position = "dodge") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) +
    labs(
      x = 'smoking group ',
      y = 'counts',
      title = 'The distribution of insurance in different smoking group ')
```

##### Chi-quare test  between insurance type and smoking

$H_0$ : insurance type and smoking are independent, there is no relationship between the two categorical variables. 

$H_1$ : insurance type and smoking are dependent, there is a relationship between the two categorical variables. 

*Test result:*

```{r}
#test_data_smoke_insurancetype = analyse_smoke_insurancetype %>% 
#  pivot_wider(
#  names_from = "insure", 
#  values_from = "n")

#test_data_smoke_insurancetype %>% knitr::kable()
#chisq.test(test_data_smoke_insurancetype[-1])

dat = 
  dataset_basic %>%
  drop_na(smoker,insure) %>% 
  mutate(smoker = factor(smoker,levels = c(1,2,3),labels = c("never","current","former")),
         everyday = factor(everyday)) %>% 
  mutate(insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid/Family Health+", "Milit/CHAMPUS/Tricare", "COBRA/Other", "Uninsured")))





# fourth method:
dat %$%
  ctable(smoker, insure,
    prop = "r", chisq = TRUE, headings = FALSE
  ) %>%
  print(
    method = "render",
    style = "rmarkdown",
    footnote = NA
  )



mosaic(~ smoker + insure,
  direction = c("v", "h"),
  data = dat,
  shade = TRUE
)
```



#### Age
##### Plot

```{r message=FALSE}
analyse_generalhealth_insurancetype = 
  dataset_basic %>%
  drop_na(generalhealth,insure) %>% 
  mutate(insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid/Family Health+", "Milit/CHAMPUS/Tricare", "COBRA/Other", "Uninsured")))%>% 
  mutate(generalhealth = factor(generalhealth,levels = c(1,2,3,4,5),labels = c("Excellent","Very good","Good", "Fair", "Poor"))) %>% 
  group_by(insure,generalhealth) %>% 
  summarize(n = n())
  
analyse_generalhealth_insurancetype %>% 
  ggplot(aes(x = generalhealth, y = n, fill = insure  )) + 
  geom_bar(stat = "identity",position = "dodge") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) +
    labs(
      x = 'health status',
      y = 'counts',
      title = 'The distribution of insurance type in groups with differnt general health status')
```

##### Chi-quare test  between insurance type and health status

$H_0$ : insurance type and health status are independent, there is no relationship between the two categorical variables. 

$H_1$ : insurance type and health status are dependent, there is a relationship between the two categorical variables. 

*Test result:*

```{r}
#test_data_smoke_insurancetype = analyse_smoke_insurancetype %>% 
#  pivot_wider(
#  names_from = "insure", 
#  values_from = "n")

#test_data_smoke_insurancetype %>% knitr::kable()
#chisq.test(test_data_smoke_insurancetype[-1])

dat = 
  dataset_basic %>%
  drop_na(generalhealth,insure) %>% 
  mutate(insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid/Family Health+", "Milit/CHAMPUS/Tricare", "COBRA/Other", "Uninsured")))%>% 
  mutate(generalhealth = factor(generalhealth,levels = c(1,2,3,4,5),labels = c("Excellent","Very good","Good", "Fair", "Poor")))





# fourth method:
dat %$%
  ctable(generalhealth, insure,
    prop = "r", chisq = TRUE, headings = FALSE
  ) %>%
  print(
    method = "render",
    style = "rmarkdown",
    footnote = NA
  )



mosaic(~ generalhealth + insure,
  direction = c("v", "h"),
  data = dat,
  shade = TRUE
)
```

#### Have children or not
##### Plot
```{r message=FALSE}
analyse_sickadvice_insurancetype = 
  dataset_basic %>%
  drop_na(sickadvice,insure) %>% 
  mutate(insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid_Family Health", "milit_champus_tricare", "COBRA_Other", "uninsured")))%>% 
  mutate(sickadvice = factor(sickadvice,levels = c(1,2,3,4,5,6,7,8,9),labels = c("private_doctor","non-retail_clinic","urgent_care_center", "hospital_ed", "retail_clinic", "alternative_health_care_provider","family_friend_self_resources","other","no_usual_place"))) %>% 
  group_by(insure,sickadvice) %>% 
  summarize(n = n())
  
analyse_sickadvice_insurancetype %>% 
  ggplot(aes(x = sickadvice, y = n, fill = insure )) + 
  geom_bar(stat = "identity",position = "dodge") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) +
    labs(
      x = 'place',
      y = 'counts',
      title = 'The distribution of insurance type in groups of people went to diffrent places when sick ')
```


##### Chi-quare test  between insurance type and sick advice place

$H_0$ : insurance type and sick advice place are independent, there is no relationship between the two categorical variables. 

$H_1$ : insurance type and sick advice place are dependent, there is a relationship between the two categorical variables. 

*Test result:*

```{r}
#test_data_smoke_insurancetype = analyse_smoke_insurancetype %>% 
#  pivot_wider(
#  names_from = "insure", 
#  values_from = "n")

#test_data_smoke_insurancetype %>% knitr::kable()
#chisq.test(test_data_smoke_insurancetype[-1])

dat = 
  dataset_basic %>%
  drop_na(sickadvice,insure) %>% 
  mutate(insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid_Family Health", "milit_champus_tricare", "COBRA_Other", "uninsured")))%>% 
  mutate(sickadvice = factor(sickadvice,levels = c(1,2,3,4,5,6,7,8,9),labels = c("private_doctor","non-retail_clinic","urgent_care_center", "hospital_ed", "retail_clinic", "alternative_health_care_provider","family_friend_self_resources","other","no_usual_place"))) 





# fourth method:
dat %$%
  ctable(sickadvice, insure,
    prop = "r", chisq = TRUE, headings = FALSE
  ) %>%
  print(
    method = "render",
    style = "rmarkdown",
    footnote = NA
  )



mosaic(~ sickadvice + insure,
  direction = c("v", "h"),
  data = dat,
  shade = TRUE
)
```
According to above chi-square test result We reject the null hypothesis and conclude that there’s at least one helath stutus group’s proportions of insurance type is different from others at 0.05 significant level.


#### BMI
##### Plot

```{r message=FALSE}
analyse_didntgetcare_insurancetype = 
  dataset_basic %>%
  drop_na(didntgetcare,insure) %>% 
  mutate(insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid_Family Health", "milit_champus_tricare", "COBRA_Other", "uninsured")))%>% 
  mutate(didntgetcare = factor(didntgetcare,levels = c(1,2),labels = c("Yes","No"))) %>% 
  group_by(insure,didntgetcare) %>% 
  summarize(n = n())
  
analyse_didntgetcare_insurancetype %>% 
  ggplot(aes(x = didntgetcare, y = n, fill = insure  )) + 
  geom_bar(stat = "identity",position = "dodge") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) +
    labs(
      x = 'medical care accessibility',
      y = 'counts',
      title = 'The distribution of insurance type in different  group of medical care accessibility')
```

##### Chi-quare test  between insurance type and medical care accessibility

$H_0$ : insurance type and medical care accessibility are independent, there is no relationship between the two categorical variables. 

$H_1$ : insurance type and medical care accessibility are dependent, there is a relationship between the two categorical variables. 

*Test result:*

```{r}
#test_data_smoke_insurancetype = analyse_smoke_insurancetype %>% 
#  pivot_wider(
#  names_from = "insure", 
#  values_from = "n")

#test_data_smoke_insurancetype %>% knitr::kable()
#chisq.test(test_data_smoke_insurancetype[-1])

dat = 
  dataset_basic %>%
  drop_na(didntgetcare,insure) %>% 
  mutate(insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid_Family Health", "milit_champus_tricare", "COBRA_Other", "uninsured")))%>% 
  mutate(didntgetcare = factor(didntgetcare,levels = c(1,2),labels = c("Yes","No")))





# fourth method:
dat %$%
  ctable(didntgetcare, insure,
    prop = "r", chisq = TRUE, headings = FALSE
  ) %>%
  print(
    method = "render",
    style = "rmarkdown",
    footnote = NA
  )



mosaic(~ didntgetcare + insure,
  direction = c("v", "h"),
  data = dat,
  shade = TRUE
)
```

