---
title: "Data exploration"
output: 
  html_document:
    code_folding: hide
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(httr)
library(haven)
library(summarytools)
library(ggstatsplot)
library(ggplot2)
library(patchwork)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


```{r message=FALSE}
ataset_basic = read_csv("data/dataset_basic.csv")

dataset_basic = read_csv("data/dataset_basic.csv") %>% 
  mutate(
         child = factor(child,levels = c(1,2),labels = c("yes","no")),
         sex = factor(sex,labels = c("male","female")),
         insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid/Family Health+", "Milit/CHAMPUS/Tricare", "COBRA/Other", "Uninsured")),
         smoker = factor(smoker,levels = c(1,2,3),labels = c("never","current","former")),
         agegroup = factor(agegroup,ordered = TRUE,labels = c("18-24yrs","25-44yrs", "45-64yrs", "65+yrs"))
         ) %>% 
  select(insure,agegroup,smoker,bmi,child,sex)
```


```{r message=FALSE}

smoke_df = read_csv("data/dataset_basic.csv") %>% 
  select(agegroup,smoker,everyday,numberperdaya,cost20cigarettes,generalhealth,poverty_group = imputed_povertygroup,bmi,child,sex) %>% 
  drop_na(smoker) %>% 
  mutate(
         smoker_cat = case_when(
           numberperdaya <= 10 ~ 1,
           (numberperdaya > 10) & (numberperdaya <= 20) ~2,
           numberperdaya > 20 ~ 3),
         smoker_cat = factor(smoker_cat,levels = c(1,2,3),labels = c("light","normal","heavy")),
         smoker = factor(smoker,levels = c(1,2,3),labels = c("never","current","former")),
         agegroup = factor(agegroup,ordered = TRUE,labels = c("18-24yrs","25-44yrs", "45-64yrs", "65+yrs")),
         generalhealth = factor(generalhealth,levels = c(5,4,3,2,1), labels = c("Poor","Fair","Good","Very good","Excellent")),
         poverty_group = factor(poverty_group,levels = c(1,2,3,4,5),labels = c("<100% FPL","100 - <200% FPL","200 - <400% FPL","400 - <600% FPL",">600% FPL")),
         everyday = factor(everyday, levels = c(1,2), labels = c("everyday smoker", "casual smoker"))
         
         )

current_smoker = smoke_df %>% filter(smoker == "current")
```


## 1. Knowledge about **insurance** in our dataset

XXXXXX

```{r message=FALSE}
insurehold = read_csv("data/dataset_basic.csv") %>% 
  drop_na() %>% 
  mutate(
  insuredgateway = factor(insuredgateway, levels = c(1,2), labels = c("yes", "no"))) %>% 
  ggplot(aes(x = insuredgateway,fill = insuredgateway))+
  geom_bar()+
  theme(legend.position = "none") +
  labs(
    x = "Hold status of insurance",
    y = "Number of participants",
    title = "Holding status of insurace among participants")
  
```


```{r }
insuretype = dataset_basic %>% 
  drop_na() %>% 
  mutate(insure = fct_infreq(insure)) %>% 
  ggplot(aes(x = insure,fill = insure))+
  geom_bar()+
  theme(legend.position = "none") +
  labs(
    x = "Insurance type",
    y = "Number of holders",
    title = "Number of holders in each insurance type") + 
  coord_flip()
  
  
```

```{r}
#| my-chunkhaha, echo = FALSE, fig.width = 10,
insurehold + insuretype
```


```{r}
insur_type <- c("Employer", "Medicare", "Milit/CHAMPUS", "Uninsured", "Self-purchased", "Medicaid/Family Health+", "COBRA/other")

insur_meaning <- c("Employer-sponsored health insurance is a health insurance selected and purchased by your employer and offered to eligible employees and their dependents.",
                   "Medicare is federal health insurance for people 65 or older, some younger people with disabilities, people with End-Stage Renal Disease",
                   "Health Care progam for military",
                   "Not insured",
                   "Insurance purchased by individual choice",
                   "Medicaid/health insurance program for adults who are aged 19 to 64 who have income too high to qualify for Medicaid",
                   "Consolidated Omnibus Budget Reconciliation Act, which gives some employees the ability to continue health insurance coverage after leaving employment, or other types of insurance.")

insur <- data.frame(insur_type, insur_meaning) %>% 
  dplyr::rename(
    "Insurance type" = insur_type,
    "Insurance type meaning" = insur_meaning
  )
knitr::kable(insur)
```

<br>

### 2. Knowledge about **smoker** in our dataset {.tabset}
#### Age distribution

##### Plot

```{r,message=FALSE}

smoke_df %>% 
  drop_na(agegroup) %>% 
  group_by(agegroup,smoker) %>% 
  summarize(
    count = n(),
  ) %>% 
  mutate(smoker = fct_reorder(smoker,count)) %>% 
  ggplot(aes(x = smoker, y = count, fill = agegroup)) + 
  geom_bar(stat = "identity",position = "dodge") +
    labs(
      x = 'Smoking status',
      y = 'Number of smokers',
      title = 'Age distribution among people in different smoking status')


```




#### Health self-evaluation

##### Plot

```{r,message=FALSE}

smokegroup_count = smoke_df %>% 
  drop_na(smoker) %>% 
  group_by(smoker) %>% 
  summarize(
    smokegroup_count = n()) %>% 
  pull()

smoke_df %>% 
  drop_na(generalhealth) %>% 
  group_by(generalhealth,smoker) %>% 
  summarize(
    count = n())%>% 
  mutate(smokegroup_count = smokegroup_count,
         health_percent = count/smokegroup_count)%>% 
  ggplot(aes(x = smoker, y = health_percent, fill = generalhealth)) + 
  geom_bar(stat = "identity",position = "dodge") +
    labs(
      x = 'Smoking status',
      y = 'Health self-evaluation \n(% in each types of smoker)',
      title = 'Self-evaluation of health conditions among people in different smoking status\n(% in each types of smoker)')
```


#### Poverty status

##### Plot

```{r,message=FALSE}
smoke_df %>% 
  drop_na(poverty_group) %>% 
  group_by(poverty_group,smoker) %>% 
  summarize(
    count = n(),
  ) %>% 
  mutate(
    smokegroup_count = smokegroup_count,
    poverty_percent = count/smokegroup_count,
    smoker = fct_reorder(smoker,count)) %>% 
  ggplot(aes(x = smoker, y = poverty_percent, fill = poverty_group)) + 
  geom_bar(stat = "identity",position = "dodge") +
    labs(
      x = 'Smoking status',
      y = 'Poverty groups',
      title = 'Poverty groups among people in different smoking status')
```


#### Current smokers

##### Smoking intensity and frequency


```{r message=FALSE, warning=FALSE}
current_smoker %>%
  ggplot(aes(x = smoker_cat, fill = everyday))+
  geom_histogram(stat="count") +
    labs(
      x = 'Type of current smokers',
      y = 'Number of smokers in each type',
      title = 'Number of current smokers of diffirent types')
```


```{r message=FALSE, warning=FALSE}
current_smoker %>% 
  ggplot(aes(x = numberperdaya, fill = everyday)) + 
  geom_histogram(alpha = .7)+
  xlim(0,30)+
  labs(
    x = "Average number of cigarattes smoked per day",
    y = "Number of smokers",
    title = "Average number of cigarattes smoked per day by current smokers"
       )+
  theme(legend.position = "bottom")
```


#### Cost on smoke

```{r message=FALSE, warning=FALSE}
  
current_smoker %>% 
  ggplot(aes(x = numberperdaya, y = cost20cigarettes,color = everyday))+
  geom_point(alpha = .5,size = 3) +
  xlim(0,50)+
  ylim(0,40)+
  #facet_grid(.~smoker_cat,scales = "free") +
    labs(
      x = 'Number of cigrattes consumed everyday',
      y = 'Cost of every 20 (one pack) cigrattes',
      title = 'Cost of every 20 (one pack) cigrattes\nfor number of cigrattes consumed everyday')


```

<br>

## 3. Explore relationship between `insurance` and target variables

<br>

### Distributions {.tabset}
#### Smoking

##### **Distribution**

```{r message=FALSE}
analyse_smoke_insurancetype = 
  dataset_basic %>%
  drop_na(smoker,insure) %>% 
  group_by(smoker,insure) %>% 
  summarize(n = n())
  
analyse_smoke_insurancetype %>% 
  ggplot(aes(x =smoker , y = n, fill = insure)) + 
  geom_bar(stat = "identity",position = "dodge") +
    labs(
      x = 'smoking group ',
      y = 'counts',
      title = 'The distribution of insurance in different smoking group ')
```

#### Age
##### **Distribution**

```{r message=FALSE}
analyse_agegroup_insurancetype = 
  dataset_basic %>%
  drop_na(agegroup,insure) %>% 
  group_by(insure,agegroup) %>% 
  summarize(n = n())
  
analyse_agegroup_insurancetype %>% 
  ggplot(aes(x = agegroup, y = n, fill = insure  )) + 
  geom_bar(stat = "identity",position = "dodge") +
    labs(
      x = 'agegroup',
      y = 'counts',
      title = 'The distribution of insurance type in differnt agegroup')
```


#### Have children or not
##### **Distribution**

```{r message=FALSE}
analyse_children_insurancetype = 
  dataset_basic %>%
  drop_na(child,insure)  %>% 
  group_by(insure,child) %>% 
  summarize(n = n())
  
analyse_children_insurancetype %>% 
  ggplot(aes(x = child, y = n, fill = insure )) + 
  geom_bar(stat = "identity",position = "dodge") +
    labs(
      x = 'child',
      y = 'counts',
      title = 'The distribution of insurance type in groups of people with or without children')
```


#### BMI
##### **Distribution**

```{r message=FALSE}
analyse_bmi_insurancetype = 
  dataset_basic %>%
  drop_na(bmi,insure) 
  
  
  
analyse_bmi_insurancetype %>% 
  ggplot(aes(x = insure, y = bmi,fill = insure)) + 
  geom_boxplot(alpha=0.6) +
    theme(legend.position="none") +
    scale_fill_brewer(palette="Dark2")+
    labs(
      x = 'insurance',
      y = 'bmi',
      title = 'The distribution of bmi in groups having different insurance type') +
  coord_flip()
```



<br>

### Statistical test {.tabset}

#### Smoking

##### **Chi-quare test between insurance type and smoking**

$H_0$ : insurance type and smoking are independent, there is no relationship between the two categorical variables. 

$H_1$ : insurance type and smoking are dependent, there is a relationship between the two categorical variables. 

*Test result:*

```{r}
dat = 
  dataset_basic %>%
  drop_na(smoker,insure) 
```

```{r}
#| my-chunk1, echo = FALSE, fig.width = 10,

# plot
ggbarstats(
  data = dat,
  x = smoker,
  y = insure,
  title = "Chi-quare test between insurance type and smoking"
) +
  labs(caption = NULL) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) 


```




#### Age
##### **Chi-quare test  between insurance type and age**

$H_0$ : insurance type and age are independent, there is no relationship between the two categorical variables. 

$H_1$ : insurance type and age are dependent, there is a relationship between the two categorical variables. 

*Test result:*

```{r}
dat = 
  dataset_basic %>%
  drop_na(agegroup,insure) 
```

```{r}
#| my-chunk2, echo = FALSE, fig.width = 10,
ggbarstats(
  data = dat,
  x = smoker,
  y = insure,
  title = "Chi-quare test result between insurance type and age"
) +
  labs(caption = NULL) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) 

```


#### Have children or not
##### **Chi-quare test  between insurance type and children**

$H_0$ : insurance type and have children or not are independent, there is no relationship between the two categorical variables. 

$H_1$ : insurance type and have children or not are dependent, there is a relationship between the two categorical variables. 

*Test result:*

```{r}
#test_data_smoke_insurancetype = analyse_smoke_insurancetype %>% 
#  pivot_wider(
#  names_from = "insure", 
#  values_from = "n")

#test_data_smoke_insurancetype %>% knitr::kable()
#chisq.test(test_data_smoke_insurancetype[-1])

dat = 
  dataset_basic %>%
  drop_na(child,insure)
```

```{r}
#| my-chunk3, echo = FALSE, fig.width = 10,

ggbarstats(
  data = dat,
  x = child,
  y = insure,
  title = "Chi-quare test between insurance type and children"
) +
  labs(caption = NULL) +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) 
```


#### BMI
##### **ANOVA test  between insurance type and BMI**

*Test result:*
```{r}
anova = aov( bmi ~insure, data = analyse_bmi_insurancetype)
anova %>% 
  broom::tidy() %>% 
  knitr::kable()
Tukey_comp = TukeyHSD(anova)

plot(Tukey_comp)
```

<br>


