---
title: "data exploration"
output: 
  html_document:
    code_folding: hide
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(httr)
library(haven)
library(summarytools)
library(ggstatsplot)
library(ggplot2)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


```{r, message=FALSE}

dataset_basic = read_csv("data/dataset_basic.csv")

smoke_df = dataset_basic %>% 
  select(agegroup,smoker,everyday,numberperdaya,cost20cigarettes,generalhealth,poverty_group = imputed_povertygroup,bmi,child,sex) %>% 
  drop_na(smoker) %>% 
  mutate(
         smoker_cat = case_when(
           numberperdaya <= 10 ~ 1,
           (numberperdaya > 10) & (numberperdaya <= 20) ~2,
           numberperdaya > 20 ~ 3),
         smoker_cat = factor(smoker_cat,levels = c(1,2,3),labels = c("light","normal","heavy")),
         smoker = factor(smoker,levels = c(1,2,3),labels = c("never","current","former")),
         agegroup = factor(agegroup,ordered = TRUE,labels = c("18-24yrs","25-44yrs", "45-64yrs", "65+yrs")),
         generalhealth = factor(generalhealth,levels = c(5,4,3,2,1), labels = c("Poor","Fair","Good","Very good","Excellent")),
         poverty_group = factor(poverty_group,levels = c(1,2,3,4,5),labels = c("<100% FPL","100 - <200% FPL","200 - <400% FPL","400 - <600% FPL",">600% FPL")),
         everyday = factor(everyday, levels = c(1,2), labels = c("everyday smoker", "casual smoker"))
         
         )

current_smoker = smoke_df %>% filter(smoker == "current")

dataset_basic = dataset_basic %>% 
  mutate(
         child = factor(child,levels = c(1,2),labels = c("yes","no")),
         sex = factor(sex,labels = c("male","female")),
         insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid/Family Health+", "Milit/CHAMPUS/Tricare", "COBRA/Other", "Uninsured")),
         smoker = factor(smoker,levels = c(1,2,3),labels = c("never","current","former")),
         agegroup = factor(agegroup,ordered = TRUE,labels = c("18-24yrs","25-44yrs", "45-64yrs", "65+yrs"))
         ) %>% 
  select(insure,agegroup,smoker,bmi,child,sex)


```


### Distributions and statistical tests of `target variables` and `insurance` {.tabset}

#### Smoking

##### **Distribution**

```{r message=FALSE} 
analyse_smoke_insurancetype = 
  dataset_basic %>%
  drop_na(smoker,insure) %>% 
  group_by(smoker,insure) %>% 
  summarize(n = n())
  
analyse_smoke_insurancetype %>% 
  ggplot(aes(x =smoker , y = n, fill = insure)) + 
  geom_bar(stat = "identity",position = "dodge") +
    labs(
      x = 'smoking group ',
      y = 'counts',
      title = 'The distribution of insurance in different smoking group ')
```

##### **Chi-quare test between insurance type and smoking**

$H_0$ : insurance type and smoking are independent, there is no relationship between the two categorical variables. 

$H_1$ : insurance type and smoking are dependent, there is a relationship between the two categorical variables. 

*Test result:*

```{r}
#test_data_smoke_insurancetype = analyse_smoke_insurancetype %>% 
#  pivot_wider(
#  names_from = "insure", 
#  values_from = "n")

#test_data_smoke_insurancetype %>% knitr::kable()
#chisq.test(test_data_smoke_insurancetype[-1])

dat = 
  dataset_basic %>%
  drop_na(smoker,insure) 

dat %$%
  ctable(smoker, insure,
    prop = "r", chisq = TRUE, headings = FALSE
  ) %>%
  print(
    method = "render",
    style = "rmarkdown",
    footnote = NA
  )


# plot
ggbarstats(
  data = dat,
  x = smoker,
  y = insure
) +
  labs(caption = NULL) # remove caption


```



#### Age
##### Plot

```{r message=FALSE}
analyse_agegroup_insurancetype = 
  dataset_basic %>%
  drop_na(agegroup,insure) %>% 
  group_by(insure,agegroup) %>% 
  summarize(n = n())
  
analyse_agegroup_insurancetype %>% 
  ggplot(aes(x = agegroup, y = n, fill = insure  )) + 
  geom_bar(stat = "identity",position = "dodge") +
    labs(
      x = 'agegroup',
      y = 'counts',
      title = 'The distribution of insurance type in differnt agegroup')
```

##### Chi-quare test  between insurance type and age

$H_0$ : insurance type and age are independent, there is no relationship between the two categorical variables. 

$H_1$ : insurance type and age are dependent, there is a relationship between the two categorical variables. 

*Test result:*

```{r}
#test_data_smoke_insurancetype = analyse_smoke_insurancetype %>% 
#  pivot_wider(
#  names_from = "insure", 
#  values_from = "n")

#test_data_smoke_insurancetype %>% knitr::kable()
#chisq.test(test_data_smoke_insurancetype[-1])

dat = 
  dataset_basic %>%
  drop_na(agegroup,insure) 

dat %$%
  ctable(agegroup, insure,
    prop = "r", chisq = TRUE, headings = FALSE
  ) %>%
  print(
    method = "render",
    style = "rmarkdown",
    footnote = NA
  )


ggbarstats(
  data = dat,
  x = smoker,
  y = insure
) +
  labs(caption = NULL) # remove caption

```

#### Have children or not
##### Plot
```{r message=FALSE}
analyse_children_insurancetype = 
  dataset_basic %>%
  drop_na(child,insure)  %>% 
  group_by(insure,child) %>% 
  summarize(n = n())
  
analyse_children_insurancetype %>% 
  ggplot(aes(x = child, y = n, fill = insure )) + 
  geom_bar(stat = "identity",position = "dodge") +
    labs(
      x = 'child',
      y = 'counts',
      title = 'The distribution of insurance type in groups of people with or without children')
```


##### Chi-quare test  between insurance type and children

$H_0$ : insurance type and have children or not are independent, there is no relationship between the two categorical variables. 

$H_1$ : insurance type and have children or not are dependent, there is a relationship between the two categorical variables. 

*Test result:*

```{r}
#test_data_smoke_insurancetype = analyse_smoke_insurancetype %>% 
#  pivot_wider(
#  names_from = "insure", 
#  values_from = "n")

#test_data_smoke_insurancetype %>% knitr::kable()
#chisq.test(test_data_smoke_insurancetype[-1])

dat = 
  dataset_basic %>%
  drop_na(child,insure)

# fourth method:
dat %$%
  ctable(child, insure,
    prop = "r", chisq = TRUE, headings = FALSE
  ) %>%
  print(
    method = "render",
    style = "rmarkdown",
    footnote = NA
  )

ggbarstats(
  data = dat,
  x = child,
  y = insure
) +
  labs(caption = NULL) # remove caption

```

#### BMI
##### Plot

```{r message=FALSE}
analyse_bmi_insurancetype = 
  dataset_basic %>%
  drop_na(bmi,insure) 
  
  
  
analyse_bmi_insurancetype %>% 
  ggplot(aes(x = insure, y = bmi,fill = insure)) + 
  geom_boxplot(alpha=0.6) +
    theme(legend.position="none") +
    scale_fill_brewer(palette="Dark2")+
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) +
    labs(
      x = 'insurance',
      y = 'bmi',
      title = 'The distribution of bmi in groups having different insurance type')
```

##### ANOVA test  between insurance type and BMI

*Test result:*
```{r}
anova = aov( bmi ~insure, data = analyse_bmi_insurancetype)
summary(anova)
Tukey_comp = TukeyHSD(anova)
Tukey_comp 
plot(Tukey_comp)
```



### Exploration on smoking status {.tabset}

#### Age distribution

##### Plot

```{r,message=FALSE}

smoke_df %>% 
  drop_na(agegroup) %>% 
  group_by(agegroup,smoker) %>% 
  summarize(
    count = n(),
  ) %>% 
  mutate(smoker = fct_reorder(smoker,count)) %>% 
  ggplot(aes(x = smoker, y = count, fill = agegroup)) + 
  geom_bar(stat = "identity",position = "dodge") +
    labs(
      x = 'Smoking status',
      y = 'Number of smokers',
      title = 'Age distribution among people in different smoking status')


```


#### Health self-evaluation

##### Plot

```{r,message=FALSE}

smokegroup_count = smoke_df %>% 
  drop_na(smoker) %>% 
  group_by(smoker) %>% 
  summarize(
    smokegroup_count = n()) %>% 
  pull()

smoke_df %>% 
  drop_na(generalhealth) %>% 
  group_by(generalhealth,smoker) %>% 
  summarize(
    count = n())%>% 
  mutate(smokegroup_count = smokegroup_count,
         health_percent = count/smokegroup_count)%>% 
  ggplot(aes(x = smoker, y = health_percent, fill = generalhealth)) + 
  geom_bar(stat = "identity",position = "dodge") +
    labs(
      x = 'Smoking status',
      y = 'Health self-evaluation \n(% in each types of smoker)',
      title = 'Self-evaluation of health conditions among people in different smoking status\n(% in each types of smoker)')
```


#### Poverty status

##### Plot

```{r,message=FALSE}
smoke_df %>% 
  drop_na(poverty_group) %>% 
  group_by(poverty_group,smoker) %>% 
  summarize(
    count = n(),
  ) %>% 
  mutate(
    smokegroup_count = smokegroup_count,
    poverty_percent = count/smokegroup_count,
    smoker = fct_reorder(smoker,count)) %>% 
  ggplot(aes(x = smoker, y = poverty_percent, fill = poverty_group)) + 
  geom_bar(stat = "identity",position = "dodge") +
    labs(
      x = 'Smoking status',
      y = 'Poverty groups',
      title = 'Poverty groups among people in different smoking status')
```


#### Current smokers

##### Smoking intensity and frequency

```{r}
current_smoker %>%
  ggplot(aes(x = smoker_cat, fill = everyday))+
  geom_histogram(stat="count") +
    labs(
      x = 'Type of current smokers',
      y = 'Number of smokers in each type',
      title = 'Number of current smokers of diffirent types')
```


```{r}
current_smoker %>% 
  ggplot(aes(x = numberperdaya, fill = everyday)) + 
  geom_histogram(alpha = .7)+
  xlim(0,30)+
  labs(
    x = "Average number of cigarattes smoked per day",
    y = "Number of smokers",
    title = "Average number of cigarattes smoked per day by current smokers"
       )+
  theme(legend.position = "bottom")
```


```{r}
current_smoker %>% 
  ggplot(aes(x = cost20cigarettes, fill = smoker_cat)) + 
  geom_histogram(alpha = .7)+
  xlim(0,20)+
  labs(
    x = "Cost on each pack (20 cigarattes)",
    y = "Number of smokers",
    title = "Cost on each pack (20 cigarattes) consumed by current smokers"
       )
  theme(legend.position = "bottom") 
  
  

current_smoker %>%
  ggplot(aes(x = smoker_cat, y = cost20cigarettes, fill = smoker_cat))+
  geom_violin() +
    labs(
      x = 'Type of smokers',
      y = 'Cost of every 20 (one pack) cigrattes',
      title = 'Cost of every 20 (one pack) cigrattes\namong different type of smokers')
  
  
current_smoker %>% 
  ggplot(aes(x = numberperdaya, y = cost20cigarettes,color = everyday))+
  geom_point(alpha = .5,size = 3) +
  xlim(0,50)+
  ylim(0,40)+
  #facet_grid(.~smoker_cat,scales = "free") +
    labs(
      x = 'Number of cigrattes consumed everyday',
      y = 'Cost of every 20 (one pack) cigrattes',
      title = 'Cost of every 20 (one pack) cigrattes\nfor number of cigrattes consumed everyday')


```






