---
title: "Mapping of NYC health center location"
output: 
  html_document:
    toc: true
    toc_float: true
---

For general public seeking assistance with signing up for health insurance or SNAP (Supplemental Nutrition Assistance Program), here is a mapping of health center locations, where health insurance enrollment and assistance with SNAP benefits (Food Stamps) are offered.

```{r load packages, echo = FALSE, message = FALSE}
library(tidyverse)
library(plotly)
```

```{r loading data, echo = FALSE}
location = read.csv("data/Primary_Care_Access_and_Planning_-_Health_Insurance_Enrollment.csv") %>% 
  janitor::clean_names() 
```

```{r plotly mapping, echo = FALSE, warning = FALSE}

# geo styling
g <- list(
  scope = 'usa',
  projection = list(type = 'albers usa'),
  showland = TRUE,
  landcolor = toRGB("gray95"),
  subunitcolor = toRGB("gray85"),
  countrycolor = toRGB("gray85"),
  countrywidth = 0.5,
  subunitwidth = 0.5
)

fig <- plot_geo(location, lat = ~latitude, lon = ~longitude)

fig <- fig %>% add_markers(
    text = ~paste(health_center, street_address, telephone_number, sep = "<br />"),
    color = ~borough, symbol = I("square"), size = I(8), hoverinfo = "text"
  )
fig <- fig %>% colorbar(title = "Borough")
fig <- fig %>% layout(
    title = 'Mapping of NYC health center location)', geo = g
  )

fig
```



Source: Department of Health and Mental Hygiene. (2018, June 20). Primary care access and planning - health insurance enrollment: NYC Open Data. Primary Care Access and Planning - Health Insurance Enrollment | NYC Open Data. Retrieved November 19, 2021, from https://data.cityofnewyork.us/Health/Primary-Care-Access-and-Planning-Health-Insurance-/gfej-by6h. 

