---
title: "Insurance cost model building"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_float: true
---

```{r, echo = FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

library(tidyverse)
library(ggpubr)
library(plotly)
library(ggcorrplot)
library(olsrr)
library(MASS) 


#knitr::opts_chunk$set(
  #fig.width = 8,
  #fig.asp = .6,
  #out.width = "90%"
  
#)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

```


```{r}
cost_df = read.csv("data/medical_cost_clean.csv") %>% 
  mutate(
    sex = as.factor(sex),
    smoker = as.factor(smoker),
    children_group = as.factor(children_group)
  )

```

<br>

<font size="3.5">In order to understand the amount of insurance price change due to smoking, along with other personal factors such age, sex, weight and whether or not having children, regression model need to be build and analysed. The model would also be useful to predict insurance price.</font> 

<br>

## A snapshot of the vairables

### Variable meanings
```{r}
var <- c("smoker", "children_group", "sex", "age", "bmi", "charges")

var_meaning <- c("0 = non-smoker, 1 = smoker", "0 = No children, 1 = Have at least 1 child", "1 = male, 2 = female", "Age range from 18 to 63", "Body mass index range from 15.96 to 53.13", "Cost of health insurance")

var_info <- data.frame(var, var_meaning) %>% 
  rename(
    "Variable name" = var,
    "Variable information" = var_meaning
  )
knitr::kable(var_info)
```

<br>

### Relationship between each variables and charge of insurance {.tabset}

#### Smoker and insurance charges

```{r}

cost_df_smoker = 
  cost_df %>%
  mutate(
    smoker = fct_recode(
      smoker,
      "Smoker" = "1",
      "Non smoker" = "0"
      ))
    
smoker_plotly <- cost_df_smoker %>% 
  plot_ly(
    x = ~smoker,
    y = ~charges,
    split = ~smoker,
    type = 'violin',
    box = list(
      visible = T
    ),
    meanline = list(
      visible = T
    )
  ) 

smoker_plotly <- smoker_plotly %>%
  layout(
    xaxis = list(
      title = "smokeing status"
    ),
    yaxis = list(
      title = "Insurance charges",
      zeroline = F
    ),
    legend=list(title=list(text='<b> Smokering Status </b>'))
  )

smoker_plotly
```

<br>

#### Having children and insurance charges

```{r}
cost_df_child = 
  cost_df %>%
  mutate(
    children_group = fct_recode(
      children_group,
      "Having at least one child" = "1",
      "No children" = "0"
      ))
    
child_plotly <- cost_df_child %>% 
  plot_ly(
    x = ~children_group,
    y = ~charges,
    split = ~children_group,
    type = 'violin',
    box = list(
      visible = T
    ),
    meanline = list(
      visible = T
    )
  ) 

child_plotly <- child_plotly %>%
  layout(
    xaxis = list(
      title = "Whether or not having children"
    ),
    yaxis = list(
      title = "Insurance charges",
      zeroline = F
    ),
    legend=list(title=list(text='<b> Whether or not having children </b>'))
  )

child_plotly
```


<br>

#### Gender and insurance charges

```{r}
cost_df_sex = 
  cost_df %>%
  mutate(
    sex = fct_recode(
      sex,
      "Male" = "1",
      "Female" = "2"
      ))
    
sex_plotly <- cost_df_sex %>% 
  plot_ly(
    x = ~sex,
    y = ~charges,
    split = ~sex,
    type = 'violin',
    box = list(
      visible = T
    ),
    meanline = list(
      visible = T
    )
  ) 

sex_plotly <- sex_plotly %>%
  layout(
    xaxis = list(
      title = "Gender"
    ),
    yaxis = list(
      title = "Insurance charges",
      zeroline = F
    ),
    legend=list(title=list(text='<b> Gender </b>'))
  )

sex_plotly
```

<br>

#### Age and insurance charges

```{r}
age_fit <- lm(charges ~ age, data = cost_df)


age_plotly <- plot_ly(
  cost_df, 
  x = ~age, 
  y = ~charges,
  color = ~age, 
  size = ~age
  ) %>% 
  add_markers(y = ~charges) %>% 
  add_lines(x = ~age, y = fitted(age_fit))


age_plotly
```

<br>

#### BMI and insurance charges

```{r}
bmi_fit <- lm(charges ~ bmi, data = cost_df)


bmi_plotly <- plot_ly(
  cost_df, 
  x = ~bmi, 
  y = ~charges,
  color = ~bmi, 
  size = ~bmi
  ) %>% 
  add_markers(y = ~charges) %>% 
  add_lines(x = ~bmi, y = fitted(bmi_fit))


bmi_plotly
```

<br>

###

From the five plots above, we can tell that smoking, age, and bmi significantly affect charge of insurance. The relationship between charges and gender or having children is unclear. Further analysis will be carried below to illustrate their relationships.

<br>

## Model selection

### Correlation plot

```{r}

cost_df_cor = cost_df %>% 
  dplyr::select(age, sex, bmi, smoker, charges, children_group)

model.matrix(~0+., data = cost_df_cor) %>% 
  cor(use="pairwise.complete.obs") %>% 
  ggcorrplot(show.diag = F, type="lower", lab=TRUE, lab_size=2)

```

There is no collinearity between any of the variables. Therefore, no need to adjust variable for model building. In addition, smoking seems to be the highest correlated variable to charges among other variables.

<br>


**Fit regression using all predictors**

```{r}
model <- lm(charges ~ smoker + children_group + sex + age + bmi, data = cost_df)
summary(model)%>% 
  broom::tidy() %>% 
  knitr::kable()
```

The model building result shows that sex is not a significant variable with p-valur > 0.05. Therefore, when we do the stepwise regression, we would likely to remove the variable of sex.

<br>

### Stepwise regression

```{r, results="hide"}
step = step(model, direction='both')
```

```{r}
summary(step)%>% 
  broom::tidy() %>% 
  knitr::kable()

```

The stepwise regression process confirms that sex is not significant variable. Therefore we remove sex variable and retain all other variables for prediction model.

<br>

### Final model with significant variables

```{r}
model1 <- lm(charges ~ smoker + children_group + age + bmi, data = cost_df)
summary(model1) %>% 
  broom::tidy() %>% 
  knitr::kable()
```

<br>

## Model diagnostic
```{r}
par(mfrow = c(2,2))
plot(model1)
```

The graph above did not show homoscedasticity of residual. Therefore, we need to transform the model.

<br>

## Model transformation

### Box-Cox Transformation

```{r}
boxcox(model1, lambda = seq(-3, 3, by = 0.25)) 
```

Since lamda = 0, try natural log transformation.

<br>

**Before transformation, remove outliers first**

```{r}
plot(model1, which = 4) # remove 544, 578, 1301

cost_df_out = cost_df[-c(544,578, 1301),]

```

The outliers are rows 544, 578, 1301. Following model building will use dataset without these outliers.

<br>

**Transform charges to ln(charges)**

```{r}

#summary(cost_df_out)
cost_df_log = cost_df_out %>% 
  mutate(
    Lncharges = log(charges)
  ) 

# fit multivariate model with log transform
model_log = lm(Lncharges ~ smoker + children_group + age + bmi, data = cost_df_log) 
summary(model_log) %>% 
  broom::tidy() %>% 
  knitr::kable()

```

Above table shows the estimates for each variables when charges is transformed to In(charges).

<br>

### Diagnostic for transformed model

```{r}
boxcox(model_log)

par(mfrow = c(2,2))
plot(model_log)
```

Lamda is near 1 now, and the residual distribution improved from un-transformed data. We conclude this model is acceptable.

<br>

## Final model

```{r}
summary(model_log) %>% 
  broom::tidy() %>% 
  knitr::kable()
```

<br>

### Decode the model estimates

Holding all other variable constant, smoking is associated with a 155% increase in insurance charges than non-smokers. Being older, higher body mass index, and having at least 1 children also increase insurance charges.

<br>

## Predict your insurance premium

Want to know your insurance cost according to our model? Go to our [Shiny dashboard](https://yiqunjin.shinyapps.io/shiny_dashboard/) and find out!

