---
title: "Model building"
output: 
  html_document:
    code_folding: hide
  
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(httr)
library(haven)
library(patchwork)
library(summarytools)

```

```{r message=FALSE}
dataset_basic = read_csv("data/dataset_basic.csv")
```


### Interested variables and smoking {.tabset}

#### insurance type

```{r message=FALSE}
analyse_smoke_insurancetype = 
  dataset_basic %>%
  drop_na(smoker,insure) %>% 
  mutate(smoker = factor(smoker,levels = c(1,2,3),labels = c("never","current","former")),
         everyday = factor(everyday)) %>% 
  mutate(insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid/Family Health+", "Milit/CHAMPUS/Tricare", "COBRA/Other", "Uninsured"))) %>% 
  group_by(smoker,insure) %>% 
  summarize(n = n())
  
analyse_smoke_insurancetype %>% 
  ggplot(aes(x = insure, y = n, fill = smoker)) + 
  geom_bar(stat = "identity",position = "dodge") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) +
    labs(
      x = 'insure type',
      y = 'counts',
      title = 'The distribution of insure in different smoking group ')
```

Anova between insurance type and smoking

$H_0$ : the variables are independent, there is no relationship between the two categorical variables. Knowing the value of one variable does not help to predict the value of the other variable

$H_1$ : the variables are dependent, there is a relationship between the two categorical variables. Knowing the value of one variable helps to predict the value of the other variable

*Test result:*
```{r}
#test_data_smoke_insurancetype = analyse_smoke_insurancetype %>% 
#  pivot_wider(
#  names_from = "insure", 
#  values_from = "n")

#test_data_smoke_insurancetype %>% knitr::kable()
```

```{r}
#chisq.test(test_data_smoke_insurancetype[-1])

dat = 
  dataset_basic %>%
  drop_na(smoker,insure) %>% 
  mutate(smoker = factor(smoker,levels = c(1,2,3),labels = c("never","current","former")),
         everyday = factor(everyday)) %>% 
  mutate(insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid/Family Health+", "Milit/CHAMPUS/Tricare", "COBRA/Other", "Uninsured")))





# fourth method:
dat %$%
  ctable(smoker, insure,
    prop = "r", chisq = TRUE, headings = FALSE
  ) %>%
  print(
    method = "render",
    style = "rmarkdown",
    footnote = NA
  )


```

#### healthstatus

```{r message=FALSE}
analyse_generalhealth_insurancetype = 
  dataset_basic %>%
  drop_na(generalhealth,insure) %>% 
  mutate(insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid/Family Health+", "Milit/CHAMPUS/Tricare", "COBRA/Other", "Uninsured")))%>% 
  mutate(generalhealth = factor(generalhealth,levels = c(1,2,3,4,5),labels = c("Excellent","Very good","Good", "Fair", "Poor"))) %>% 
  group_by(insure,generalhealth) %>% 
  summarize(n = n())
  
analyse_generalhealth_insurancetype %>% 
  ggplot(aes(x = insure, y = n, fill = generalhealth )) + 
  geom_bar(stat = "identity",position = "dodge") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) +
    labs(
      x = 'generalhealth',
      y = 'counts',
      title = 'The distribution of generalhealth in different smoking group ')
```

```{r}
test_data_generalhealth_insurancetype = analyse_generalhealth_insurancetype %>% 
  pivot_wider(
  names_from = "insure", 
  values_from = "n")
```

```{r}
test_generalhealth = chisq.test(test_data_generalhealth_insurancetype[-1])

```

#### place to go when sick?

```{r message=FALSE}
analyse_sickadvice_insurancetype = 
  dataset_basic %>%
  drop_na(sickadvice,insure) %>% 
  mutate(insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid_Family Health", "milit_champus_tricare", "COBRA_Other", "uninsured")))%>% 
  mutate(sickadvice = factor(sickadvice,levels = c(1,2,3,4,5,6,7,8,9),labels = c("private_doctor","non-retail_clinic","urgent_care_center", "hospital_ed", "retail_clinic", "alternative_health_care_provider","family_friend_self_resources","other","no_usual_place"))) %>% 
  group_by(insure,sickadvice) %>% 
  summarize(n = n())
  
analyse_sickadvice_insurancetype %>% 
  ggplot(aes(x = insure, y = n, fill = sickadvice )) + 
  geom_bar(stat = "identity",position = "dodge") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) +
    labs(
      x = 'sickadvice',
      y = 'counts',
      title = 'The distribution of sickadvice in different smoking group ')
```

```{r}
test_data_sickadvice_insurancetype = analyse_sickadvice_insurancetype %>% 
  pivot_wider(
  names_from = "insure", 
  values_from = "n") %>% 
  mutate(milit_champus_tricare = replace_na(milit_champus_tricare, 0))
```

```{r}
test_generalhealth = chisq.test(test_data_sickadvice_insurancetype[-1])

```
According to above chi-square test result We reject the null hypothesis and conclude that there’s at least one helath stutus group’s proportions of insurance type is different from others at 0.05 significant level.


#### medical care accessibility?

```{r message=FALSE}
analyse_didntgetcare_insurancetype = 
  dataset_basic %>%
  drop_na(didntgetcare,insure) %>% 
  mutate(insure = factor(insure,levels = c(1,2,3,4,5,6,7),labels = c("Employer","Self-purchase","Medicare", "Medicaid_Family Health", "milit_champus_tricare", "COBRA_Other", "uninsured")))%>% 
  mutate(didntgetcare = factor(didntgetcare,levels = c(1,2),labels = c("Yes","No"))) %>% 
  group_by(insure,didntgetcare) %>% 
  summarize(n = n())
  
analyse_didntgetcare_insurancetype %>% 
  ggplot(aes(x = insure, y = n, fill = didntgetcare )) + 
  geom_bar(stat = "identity",position = "dodge") +
  theme(
    axis.text.x = element_text(angle = 90, vjust = .5, hjust = 1)
  ) +
    labs(
      x = 'didntgetcare',
      y = 'counts',
      title = 'The distribution of didntgetcare in different smoking group ')
```

```{r}
test_data_didntgetcare_insurancetype = analyse_didntgetcare_insurancetype %>% 
  pivot_wider(
  names_from = "insure", 
  values_from = "n") 
  #mutate(milit_champus_tricare = replace_na(milit_champus_tricare, 0))
```

```{r}
test_generalhealth = chisq.test(test_data_didntgetcare_insurancetype[-1])

```